Omix

### 2019-04-08 omix/omix.Rmd
reorg omix files

```{r setup, include=FALSE}
set.seed(1234)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# num_cores <- parallel::detectCores()
options(mc.cores = parallel::detectCores(), expressions = 5e5)
```

@@

DepMap 19Q1 update fixed almost all of the inconsistencies of the Primary.Disease column, though there were come capitalization inconsistencies.
```{r}
ccl_info <- read.delim("./data_munging/19Q1/DepMap-2019q1-celllines_v2_20190305.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_info$source <- "CCLE"
ccl_info$Cell_line <- strsplit(as.character(ccl_info$CCLE_Name), split = "_")
ccl_info$Cell_line <- as.character(lapply(ccl_info$Cell_line, function(x) ifelse("[MERGED" %in% x || "NORMAL" %in% x, NA, x[1])))
ccl_info$Cell_line[ccl_info$Cell_line == "NA"] <- NA
ccl_info$source_histology <- tolower(gsub("^[a-z0-9]*_", "", ccl_info$CCLE_Name, ignore.case = TRUE))
ccl_info$source_histology <- ifelse(ccl_info$CCLE_Name == ccl_info$CCLE_Name, NA, ccl_info$source_histology)

# This CCL had a histology of "primary": Looking at the DepMap reference, this is an immortalize epithelial line from the CNS (https://depmap.org/portal/cell_line/NHAHTDD_PRIMARY), so I chanegd the histology accordingly
ccl_info$source_histology[ccl_info$Cell_line == "NHAHTDD"] <- "central_nervous_system"
# Convert Primary.Disease entries to all lower case to fix factorization issue
ccl_info$Primary.Disease <- as.factor(tolower(ccl_info$Primary.Disease))
# Harmonize ID column name and rename CCLE_Name column to fit with dataframe style
colnames(ccl_info)[colnames(ccl_info) == "CCLE_Name"] <- "CCLE_ccl_name"
colnames(ccl_info)[colnames(ccl_info) == "DepMap_ID"] <- "Broad_ID"
```

Adding TCGA labels:
```{r, eval=FALSE}
tcga_hist_converter <- read.xlsx("./data_munging/TCGA/cell_line_annotation.xlsx", sheet = 1, startRow = 1, colNames = TRUE, rowNames = FALSE)
tcga_hist_converter <- add_column(.data = tcga_hist_converter, TCGA_manual = NA, .after = "TCGA_combine")
tcga_hist_converter <- tcga_hist_converter[with(tcga_hist_converter, order(primary_DepMap, subtype_DepMap, primary_site_CTRP, primary_hist_CTRP, subtype_CTRP, tissue1_GDSC)),]

# For manual curation of TCGA labels
# ccls_for_manual_tcga_curation <- filter(tcga_hist_converter, is.na(TCGA_combine))
# ccls_for_manual_tcga_curation <- ccls_for_manual_tcga_curation[with(ccls_for_manual_tcga_curation, order(primary_DepMap, subtype_DepMap, primary_site_CTRP, primary_hist_CTRP, subtype_CTRP, tissue1_GDSC)),]
# write.csv(ccls_for_manual_tcga_curation, "./data_munging/TCGA/cell_line_annotation_noTCGA.csv")
tcga_hist_converter <- filter(tcga_hist_converter, !is.na(TCGA_combine))
ccls_curated <- read.delim("./data_munging/TCGA/cell_line_annotation_noTCGA_curated.csv", sep = ",", row.names = 1)
tcga_histsconverter <- bind_rows(tcga_hist_converter, ccls_curated)
tcga_hist_converter$TCGA_combine <- with(tcga_hist_converter, ifelse(is.na(TCGA_combine), as.character(TCGA_manual), TCGA_combine))
# write.csv(tcga_hist_converter, "./data_munging/TCGA/cell_line_annotation_curated_sally_20190404.csv")
```

## Omics

@@

Mutation calls (coding region, germline filtered):
```{r, eval=FALSE}
maf_raw <- read.delim("./data_munging/19Q1/depmap_19Q1_mutation_calls_20190305.csv.gz", header = TRUE, sep = ",", row.names = 1)
# Harmonize ID column name
colnames(maf_raw)[colnames(maf_raw) == "DepMap_ID"] <- "Broad_ID"
# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "Broad_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1", "Genome_Change"))
maf_df$Reference_Allele <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1 <- nchar(as.character(maf_df$Tumor_Seq_Allele1))
maf_df <- maf_df[maf_df$Variant_Classification != "Silent",]
# Add Mutation_Status column
maf_df$Mutation_Status_Nonsilent <- "Mutant"

saveRDS(maf_df, "./data_munging/19Q1/maf_df_19Q1.rds", compress = "xz")
```

DepMap WES CN Data:
```{r, eval=FALSE}
cn <- read.delim("./data_munging/19Q1/public_19Q1_gene_cn_20190305.csv.gz", sep = ",", check.names = FALSE, header = TRUE)
# Remove Entrez gene IDs from colnames, format is "Hugo_Symbol (Entrez ID)"
colnames(cn) <- gsub(" .*", "", colnames(cn))
colnames(cn)[1] <- "Broad_ID"
# Melt
cn_melt <- melt(data = cn, id.vars = "Broad_ID", measure.vars = colnames(cn[2:ncol(cn)]), variable.name = "Hugo_Symbol", value.name = "CN")
saveRDS(cn_melt, "./data_munging/19Q1/cn_melt_19Q1.rds", compress = "xz")
```

19Q1, CCLE RNAseq gene expression data (log2(TPM+1)):
```{r, eval=FALSE}
ge_19Q1 <- read.delim("./../htscreens_giant_files/CCLE_depMap_19Q1_TPM_20190305.csv.gz", header = TRUE, sep = ",", check.names = FALSE)
# Remove Entrez gene IDs from colnames, format is "Hugo_Symbol (Entrez ID)"
colnames(ge_19Q1) <- gsub(" .*", "", colnames(ge_19Q1))
colnames(ge_19Q1)[1] <- "Broad_ID"

# For Denise's GCN
pgx_ccle_ge <- merge(select(ccl_info, source, source_ccl_name, source_histology, sample_ID, CCLE_ccl_name, Broad_ID), ge_19Q1, by = "Broad_ID", all = TRUE)
saveRDS(pgx_ccle_ge, "./data_munging/19Q1/pgx_ccle_ge.rds", compress = "xz")

# Melt
ge_melt_19Q1 <- melt(ge_19Q1, id.vars = "Broad_ID", measure.vars = colnames(ge_19Q1)[2:ncol(ge_19Q1)], variable.name = "Hugo_Symbol", value.name = "TPM")
# ge_melt_19Q1 <- unique(ge_melt_19Q1_all)
saveRDS(ge_melt_19Q1, "./data_munging/19Q1/ge_melt_19Q1.rds", compress = "xz")
```

Load Rds:
```{r}
maf_df <- readRDS("./data_munging/19Q1/maf_df_19Q1.rds")
cn_melt <- readRDS("./data_munging/19Q1/cn_melt_19Q1.rds")
ge_melt <- readRDS("./data_munging/19Q1/ge_melt_19Q1.rds")
```

@@

```{r, eval=FALSE}
# Sample info from figshare
crispr_meta <- read.delim("./data_munging/19Q1/sample_info_20190306.csv", sep = ",", header = TRUE, na.strings = c("", NA))
colnames(crispr_meta)[colnames(crispr_meta) == "CCLE_name"] <- "CCLE_Name"

crispr <- read.delim("./data_munging/19Q1/gene_effect_corrected_20190305.csv.gz", sep = ",", check.names = FALSE, header = TRUE)
```

# Pozdeyev et al. (2016) summary

@@

```{r}
ccl_histologies <- select(ccl_info, Cell_line, source_histology, CCLE_ccl_name, Broad_ID) %>% drop_na(Cell_line)
colnames(ccl_histologies)[colnames(ccl_histologies) == "source_histology"] <- "Histology"

# Account for cell line aliases and missing histologies
auc_na_histologies <- read.delim("./data_munging/pozdeyev2016/auc_na_histologies.csv", sep = ",", row.names = 1)

auc_na_histologies$accession_id <- NULL
auc_na_histologies$Notes <- NULL
auc_na_histologies$Histology <- ifelse(auc_na_histologies$Histology == "", NA, as.character(auc_na_histologies$Histology))
auc_na_histologies$CCLE_ccl_name <- ifelse(auc_na_histologies$CCLE_ccl_name  == "", NA, as.character(auc_na_histologies$CCLE_ccl_name ))
auc_na_histologies$Broad_ID <- ifelse(auc_na_histologies$Broad_ID == "", NA, as.character(auc_na_histologies$Broad_ID))

@@

```{r, eval=FALSE}
ccle_poz <- read.delim("./data_munging/pozdeyev2016/CCLE_drug_response_metrics.csv.gz", sep = ",")

# ccle_test <- merge(select(ccle_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "CCLE"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

ctrp_poz <- read.delim("./../htscreens_giant_files/pozdeyev2016/CTRP_drug_response_metrics.csv.gz", sep = ",")
# ctrp_test <- merge(select(ctrp_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "CTRP"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

gdsc_poz <- read.delim("./data_munging/pozdeyev2016/GDSC_drug_response_metrics.csv.gz", sep = ",")
# gdsc_test <- merge(select(gdsc_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "GDSC"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

@@

auc$WasDrugScreened <- 1
ccls_not_drugscreened <- filter(ccl_histologies, Cell_line %in% setdiff(ccl_histologies$Cell_line, auc_wide$Cell_line))
ccls_not_drugscreened$WasDrugScreened <- 0

@@

# Identify cell lines without histology indications. Write to file to edit for merging back to ccl_histologies file
# test <- filter(auc, is.na(Histology))[, 1:10]
# auc_na_hists <- unique(filter(auc, is.na(Histology))$Cell_linet)
# write.csv(auc_na_hists, "./data_munging/pozdeyev2016/auc_na_histologies.csv")

# Create merged dataset with sample_ids in the rows and drugs across the columns
auc_wide <- dcast(auc, Cell_line + Histology + CCLE_ccl_name + Broad_ID + WasDrugScreened ~ Drug_name_Database, value.var = "AUC_IC50", fun.aggregate = mean, fill = 1000000)
auc_wide[auc_wide == 1000000] <- ""
auc_wide <- auc_wide %>% group_by(Cell_line, Histology, CCLE_ccl_name, Broad_ID, WasDrugScreened) %>% summarise_all(funs(trimws(paste(., collapse = ""))))
auc_wide[auc_wide == ""] <- NA

@@

# auc_all[, -c(1:4)] <- apply(auc_all[, -c(1:4)], MARGIN = 2, FUN = as.numeric)
auc_all <- subset(auc_all, !(WasDrugScreened == 0 & WasCRISPRScreened == 0 & HasMutationData == 0 & HasGeneExpressionData == 0 & HasCopyNumberData == 0))
# write.csv.gz(auc_all, file = "./data_munging/pozdeyev2016/merged_auc_v20190402.4.csv.gz", row.names = FALSE, na = "NA")

@@

```{r}
auc_all <- read.delim("./data_munging/pozdeyev2016/merged_auc_v20190402.4.csv.gz", sep = ",", check.names = FALSE)

@@

## Plots

```{r, eval=FALSE}
futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_pgx_crispr_overlap <- venn.diagram(list("CRISPR" = unique(crispr_ccls$Cell_line), "Drug Screen" = unique(pgx_ccls$Cell_line), "Unscreened" = unique(unscreened_ccls$Cell_line)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened via small-molecule and CRISPR", main.fontface = "bold", lty = rep("blank", 3), fill = c("mediumpurple1", "turquoise4", "slategray3"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./gcn_denise/pozdeyev_unmatched/venn_pgx_crispr_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_pgx_crispr_overlap)
dev.off()

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_ccl_overlap <- venn.diagram(list("CCLE" = unique(ccle_ccls$Cell_line), "CTRP" = unique(ctrp_ccls$Cell_line), "GDSC" = unique(gdsc_ccls$Cell_line)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened in each dataset", main.fontface = "bold", lty = rep("blank", 3), fill = c("springgreen4", "steelblue4", "turquoise4"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./gcn_denise/pozdeyev_unmatched/venn_ccl_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_ccl_overlap)
dev.off()

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_omics_overlap <- venn.diagram(list("Gene Expression" = unique(ge_ccls$Cell_line), "Copy Number" = unique(cn_ccls$Cell_line), "Mutation" = unique(mut_ccls$Cell_line)), resolution = 5000, main = "Overlapping omic annotations in cancer cell lines (CCLs)", main.fontface = "bold", lty = rep("blank", 3), fill = c("yellow2", "orange", "tomato"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./gcn_denise/pozdeyev_unmatched/venn_omics_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_omics_overlap)
dev.off()

bar_ccl_per_hist <- ggplot(data = unique(select(auc_all, Cell_line, Histology))) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  scale_y_continuous(breaks = seq(0, 300, by = 25), labels = seq(0, 300, by = 25)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology across all cell lines") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_ccl_per_hist
ggsave("./gcn_denise/pozdeyev_unmatched/barplot_ccl_per_histology.png", bar_ccl_per_hist, device = "png", dpi = 450, width = 12, height = 6, units = "in")

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_drugs_overlap <- venn.diagram(list("CCLE" = unique(ccle_poz$Drug_name), "CTRP" = unique(ctrp_poz$Drug_name), "GDSC" = unique(gdsc_poz$Drug_name)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened in each dataset", main.fontface = "bold", lty = rep("blank", 3), fill = c("springgreen4", "steelblue4", "turquoise4"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./gcn_denise/pozdeyev_unmatched/venn_drugs_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_drugs_overlap)
dev.off()

bar_ccl_per_hist_per_db <- ggplot(data = unique(select(auc, Database, Cell_line, Histology))) +
  facet_wrap(~ Database) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  # coord_cartesian(ylim = c(0, 130)) +
  scale_y_continuous(breaks = seq(0, 160, by = 20), labels = seq(0, 160, by = 20)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology, grouped by database") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_ccl_per_hist_per_db
ggsave("./gcn_denise/pozdeyev_unmatched/barplot_ccl_per_histology_per_db.png", bar_ccl_per_hist_per_db, device = "png", dpi = 450, width = 12, height = 8, units = "in")

bar_ccl_per_hist <- ggplot(data = unique(select(auc_all, Cell_line, Histology))) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  scale_y_continuous(breaks = seq(0, 220, by = 20), labels = seq(0, 220, by = 20)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology across all annotated cell lines") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_ccl_per_hist
ggsave("./gcn_denise/pozdeyev_unmatched/barplot_ccl_per_histology.png", bar_ccl_per_hist, device = "png", dpi = 450, width = 12, height = 8, units = "in")

>>>>

```{r setup, include=FALSE}
set.seed(1234)
setwd("/Users/claris01/Box\ Sync/SallyProjects/htscreens/omix")

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# num_cores <- parallel::detectCores()
options(mc.cores = parallel::detectCores(), expressions = 5e5)
```

@@

DepMap 19Q1 update fixed almost all of the inconsistencies of the Primary.Disease column, though there were come capitalization inconsistencies.
```{r}
ccl_info <- read.delim("./data/DepMap/DepMap-2019q1-celllines_v2_20190305.csv", sep = ",", header = TRUE, na.strings = c("", NA))

ccl_info <- subset(ccl_info, Primary.Disease != "unknown" & Primary.Disease != "primary cells" & !grepl("MERGED", CCLE_Name, fixed = TRUE))
ccl_info$Cell_line <- as.character(lapply(ccl_info$CCLE_Name, function(x) strsplit(as.character(x), split = "_")[[1]][1]))
ccl_info$Histology <- tolower(gsub("^[a-z0-9]*_", "", ccl_info$CCLE_Name, ignore.case = TRUE))

# Convert Primary.Disease entries to all lower case to fix factorization issue
ccl_info$Primary.Disease <- as.factor(tolower(ccl_info$Primary.Disease))

# Harmonize ID column name
colnames(ccl_info)[colnames(ccl_info) == "DepMap_ID"] <- "Broad_ID"

# Remove two rows that contain cell lines with two different histologies
# The histologies that are kept are verified through matching in Cellosaurus and Cell Model Passports
# KMH2, haematopoietic_and_lymphoid: https://web.expasy.org/cellosaurus/CVCL_1330, https://cellmodelpassports.sanger.ac.uk/passports/SIDM01018
# MS1, lung: https://web.expasy.org/cellosaurus/CVCL_1429, https://cellmodelpassports.sanger.ac.uk/passports/SIDM00245
ccl_info <- subset(ccl_info, CCLE_Name != "KMH2_THYROID" & CCLE_Name != "MS1_SKIN")
```

Adding TCGA labels:
```{r}
tcga_hist_converter <- read.xlsx("./data/TCGA/Histology_DepMap_Xena_04052019.xlsx", sheet = 3, startRow = 1, colNames = TRUE, rowNames = FALSE)
tcga_hist_converter$data_source <- NULL

tcga_hist_converter_depmap <- filter(tcga_hist_converter, grepl("^ACH", sample_ID))
tcga_hist_converter_tcga <- filter(tcga_hist_converter, grepl("^TCGA", sample_ID))
colnames(tcga_hist_converter_depmap) <- c("Broad_ID", "TCGA_classification")
colnames(tcga_hist_converter_tcga) <- c("sample", "TCGA_classification")

ccl_info <- merge(ccl_info, tcga_hist_converter_depmap, by = "Broad_ID", all.x = TRUE)
```

@@

Mutation calls (coding region, germline filtered):
```{r, eval=FALSE}
maf_raw <- read.delim("./../data_munging/19Q1/depmap_19Q1_mutation_calls_20190305.csv.gz", header = TRUE, sep = ",", row.names = 1)
# Harmonize ID column name
colnames(maf_raw)[colnames(maf_raw) == "DepMap_ID"] <- "Broad_ID"
# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "Broad_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1", "Genome_Change"))
maf_df$Reference_Allele <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1 <- nchar(as.character(maf_df$Tumor_Seq_Allele1))
maf_df <- maf_df[maf_df$Variant_Classification != "Silent",]
# Add Mutation_Status column
maf_df$Mutation_Status_Nonsilent <- "Mutant"

saveRDS(maf_df, "./../data_munging/19Q1/maf_df_19Q1.rds", compress = "xz")
```

DepMap WES CN Data:
```{r, eval=FALSE}
cn <- read.delim("./../data_munging/19Q1/public_19Q1_gene_cn_20190305.csv.gz", sep = ",", check.names = FALSE, header = TRUE)
# Remove Entrez gene IDs from colnames, format is "Hugo_Symbol (Entrez ID)"
colnames(cn) <- gsub(" .*", "", colnames(cn))
colnames(cn)[1] <- "Broad_ID"
# Melt
cn_melt <- melt(data = cn, id.vars = "Broad_ID", measure.vars = colnames(cn[2:ncol(cn)]), variable.name = "Hugo_Symbol", value.name = "CN")
saveRDS(cn_melt, "./../data_munging/19Q1/cn_melt_19Q1.rds", compress = "xz")
```

19Q1, CCLE RNAseq gene expression data (log2(TPM+1)):
```{r, eval=FALSE}
ge_19Q1 <- read.delim("./../../htscreens_giant_files/CCLE_depMap_19Q1_TPM_20190305.csv.gz", header = TRUE, sep = ",", check.names = FALSE)
# Remove Entrez gene IDs from colnames, format is "Hugo_Symbol (Entrez ID)"
colnames(ge_19Q1) <- gsub(" .*", "", colnames(ge_19Q1))
colnames(ge_19Q1)[1] <- "Broad_ID"

# For Denise's GCN
# pgx_ccle_ge <- merge(select(ccl_info, source, source_ccl_name, source_histology, sample_ID, CCLE_ccl_name, Broad_ID), ge_19Q1, by = "Broad_ID", all = TRUE)
# saveRDS(pgx_ccle_ge, "./data_munging/19Q1/pgx_ccle_ge.rds", compress = "xz")

# Melt
ge_melt_19Q1 <- melt(ge_19Q1, id.vars = "Broad_ID", measure.vars = colnames(ge_19Q1)[2:ncol(ge_19Q1)], variable.name = "Hugo_Symbol", value.name = "TPM")
# ge_melt_19Q1 <- unique(ge_melt_19Q1_all)
saveRDS(ge_melt_19Q1, "./../data_munging/19Q1/ge_melt_19Q1.rds", compress = "xz")
```

Load omics RDS:
```{r}
maf_df <- readRDS("./../data_munging/19Q1/maf_df_19Q1.rds")
cn_melt <- readRDS("./../data_munging/19Q1/cn_melt_19Q1.rds")
ge_melt <- readRDS("./../data_munging/19Q1/ge_melt_19Q1.rds")
```

@@

```{r, eval=TRUE}
# Sample info from figshare
crispr_meta <- read.delim("./data/DepMap/sample_info_20190306.csv", sep = ",", header = TRUE, na.strings = c("", NA))
colnames(crispr_meta)[colnames(crispr_meta) == "CCLE_name"] <- "CCLE_Name"

# crispr <- read.delim("./../data_munging/19Q1/gene_effect_corrected_20190305.csv.gz", sep = ",", check.names = FALSE, header = TRUE)
```

# POZDEYEV ET AL (2016) SUMMARY

@@

```{r}
ccl_histologies <- select(ccl_info, Cell_line, Histology, TCGA_classification, CCLE_Name, Broad_ID) %>% drop_na(Cell_line)

# Account for cell line aliases and missing histologies
auc_na_histologies <- read.delim("./data/pozdeyev2016/auc_na_histologies.csv", sep = ",", row.names = 1)
auc_na_histologies$accession_id <- NULL
auc_na_histologies$Notes <- NULL

auc_na_histologies$Histology <- ifelse(auc_na_histologies$Histology == "", NA, as.character(auc_na_histologies$Histology))
auc_na_histologies$CCLE_Name <- ifelse(auc_na_histologies$CCLE_Name  == "", NA, as.character(auc_na_histologies$CCLE_Name ))
auc_na_histologies$Broad_ID <- ifelse(auc_na_histologies$Broad_ID == "", NA, as.character(auc_na_histologies$Broad_ID))

@@

```{r, eval=FALSE}
ccle_poz <- read.delim("./data/pozdeyev2016/CCLE_drug_response_metrics.csv.gz", sep = ",")
# ccle_test <- merge(select(ccle_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "CCLE"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

ctrp_poz <- read.delim("./../../htscreens_giant_files/omix/pozdeyev2016/CTRP_drug_response_metrics.csv.gz", sep = ",")
# ctrp_test <- merge(select(ctrp_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "CTRP"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

gdsc_poz <- read.delim("./data/pozdeyev2016/GDSC_drug_response_metrics.csv.gz", sep = ",")
# gdsc_test <- merge(select(gdsc_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "GDSC"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

@@

auc$WasDrugScreened <- 1
ccls_not_drugscreened <- filter(ccl_histologies, Cell_line %in% setdiff(ccl_histologies$Cell_line, auc$Cell_line))
ccls_not_drugscreened$WasDrugScreened <- 0

@@

# Identify cell lines without histology indications. Write to file to edit for merging back to ccl_histologies file
# test <- filter(auc, is.na(Histology))[, 1:10]
# auc_na_hists <- unique(filter(auc, is.na(Histology))$Cell_line)
# write.csv(auc_na_hists, "./data/pozdeyev2016/auc_na_histologies.csv")

# Create merged dataset with sample_ids in the rows and drugs across the columns
auc_wide <- dcast(auc, Cell_line + Histology + TCGA_classification + CCLE_Name + Broad_ID + WasDrugScreened ~ Drug_name_Database, value.var = "AUC_IC50", fun.aggregate = mean, fill = 1000000)
auc_wide[auc_wide == 1000000] <- ""
auc_wide[auc_wide == ""] <- NA

@@

# auc_all[, -c(1:4)] <- apply(auc_all[, -c(1:4)], MARGIN = 2, FUN = as.numeric)
auc_all <- subset(auc_all, !(WasDrugScreened == 0 & WasCRISPRScreened == 0 & HasMutationData == 0 & HasGeneExpressionData == 0 & HasCopyNumberData == 0))
write.csv.gz(auc_all, file = "./data/pozdeyev2016/merged_auc_v20190408.5.csv.gz", row.names = FALSE, na = "NA")

@@

```{r}
auc_all <- read.delim("./data/pozdeyev2016/merged_auc_v20190408.5.csv.gz", sep = ",", check.names = FALSE)

@@

## Exploratory plots

```{r, eval=FALSE}
futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_pgx_crispr_overlap <- venn.diagram(list("CRISPR" = unique(crispr_ccls$Cell_line), "Drug Screen" = unique(pgx_ccls$Cell_line), "Unscreened" = unique(unscreened_ccls$Cell_line)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened via small-molecule and CRISPR", main.fontface = "bold", lty = rep("blank", 3), fill = c("mediumpurple1", "turquoise4", "slategray3"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./pozdeyev_unmatched/venn_pgx_crispr_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_pgx_crispr_overlap)
dev.off()

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_ccl_overlap <- venn.diagram(list("CCLE" = unique(ccle_ccls$Cell_line), "CTRP" = unique(ctrp_ccls$Cell_line), "GDSC" = unique(gdsc_ccls$Cell_line)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened in each dataset", main.fontface = "bold", lty = rep("blank", 3), fill = c("springgreen4", "steelblue4", "turquoise4"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./pozdeyev_unmatched/venn_ccl_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_ccl_overlap)
dev.off()

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_omics_overlap <- venn.diagram(list("Gene Expression" = unique(ge_ccls$Cell_line), "Copy Number" = unique(cn_ccls$Cell_line), "Mutation" = unique(mut_ccls$Cell_line)), resolution = 5000, main = "Overlapping omic annotations in cancer cell lines (CCLs)", main.fontface = "bold", lty = rep("blank", 3), fill = c("yellow2", "orange", "tomato"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./pozdeyev_unmatched/venn_omics_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_omics_overlap)
dev.off()

bar_ccl_per_hist <- ggplot(data = unique(select(auc_all, Cell_line, Histology))) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  scale_y_continuous(breaks = seq(0, 300, by = 25), labels = seq(0, 300, by = 25)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology across all cell lines") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_ccl_per_hist
ggsave("./pozdeyev_unmatched/barplot_ccl_per_histology.png", bar_ccl_per_hist, device = "png", dpi = 450, width = 12, height = 6, units = "in")

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_drugs_overlap <- venn.diagram(list("CCLE" = unique(ccle_poz$Drug_name), "CTRP" = unique(ctrp_poz$Drug_name), "GDSC" = unique(gdsc_poz$Drug_name)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened in each dataset", main.fontface = "bold", lty = rep("blank", 3), fill = c("springgreen4", "steelblue4", "turquoise4"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./pozdeyev_unmatched/venn_drugs_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_drugs_overlap)
dev.off()

bar_ccl_per_hist_per_db <- ggplot(data = unique(select(auc, Database, Cell_line, Histology))) +
  facet_wrap(~ Database) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  # coord_cartesian(ylim = c(0, 130)) +
  scale_y_continuous(breaks = seq(0, 160, by = 20), labels = seq(0, 160, by = 20)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology, grouped by database") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_ccl_per_hist_per_db
ggsave("./pozdeyev_unmatched/barplot_ccl_per_histology_per_db.png", bar_ccl_per_hist_per_db, device = "png", dpi = 450, width = 12, height = 8, units = "in")

bar_ccl_per_hist <- ggplot(data = unique(select(auc_all, Cell_line, Histology))) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  scale_y_continuous(breaks = seq(0, 375, by = 25), labels = seq(0, 375, by = 25)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology across all annotated cell lines") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_ccl_per_hist
ggsave("./pozdeyev_unmatched/barplot_ccl_per_histology.png", bar_ccl_per_hist, device = "png", dpi = 450, width = 12, height = 8, units = "in")

histo_aucic50_per_db <- ggplot(data = auc) +
  facet_wrap(~ Database, scales = "free_y") +
  geom_histogram(aes(x = AUC_IC50, fill = Database), color = "black", alpha = 0.7, binwidth = 0.025, boundary = 0, closed = "left") +
  scale_fill_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  guides(fill = "none") +
  labs(x = "AUC IC50", y = "Frequency", title = "Distributions of AUC IC50 values for all drugs in each dataset (unmatched)")
histo_aucic50_per_db
ggsave("./pozdeyev_unmatched/histogram_aucic50_per_db.png", histo_aucic50_per_db, device = "png", dpi = 450, width = 12, height = 6, units = "in")
```

## README text
